Improving real-time trend estimates using local parametrisation of polynomial regression filters

### 2024-04-25 paper/article_jos.qmd
update article

In @sec-propMM, we describe the general properties of moving averages and the associated quality criteria.
This allows us to understand the foundations behind the construction of moving averages from local polynomial regressions, as well as those behind the two extensions proposed in this article (@sec-nonparamreg).
In a final section (@sec-comparison), all these methods are compared empirically on simulated and real series.

@@

The basic assumption  used in  time series decomposition methods is that the input observed time series, $y_t$ can be decomposed into a signal component $\mu_t$ and a an erratic component $\varepsilon_t$ (called irregular):
$$
f(y_t)=\mu_t+\varepsilon_t
$$
where $f$ denotes some appropriate transformation (usually logarithmic or no transformation).
To simplify further notations, $y_t$ will denote the transformed observed series.
The noise component $\varepsilon_t$ is usually assumed to be a white noise.
Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $\mu_t$ represents the trend and cycle components, here estimated jointly and called trend-cycle $TC_t$.
This component can be locally approximated by a polynomial degree (see @sec-nonparamreg):
$$
\mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
$$
with $\xi_t$ a stochastic process uncorrelated with $\varepsilon_t$.
Event if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], an usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
Thus, the trend-cycle $TC_t$ is considered as deterministic and modelled as a polynomial trend of degree $d$.
The coefficients $(\beta_0,\dots,\beta_k)$ can be estimated by weighted least squares.
The estimate $\hat \beta_0$ provides the trend-cycle estimate $\hat TC_t$ and it can be shown that this equivalent to apply of moving average (see @sec-nonparamreg).
The goal of this paper is to present two new methods to separate the trend-cycle component from the noise component in real-time (i.e. when no future point is known), using moving averages build on local polynomial smoothing techniques.

@@

Appendix [-@sec-an-cof] shows the coefficients, gain and phase shift functions of the four asymmetric filters.

@@

The series associa

@@

@@

```{r, echo=FALSE, out.width="100%"}
#| label: fig-graphstpsimul
#| fig-cap: Distribution of phase shifts on simulated series.
img <- sprintf("img/simulations/phase_shift_simul.%s", "pdf")
knitr::include_graphics(img)
```

@@

# Discussion
## Outliers and COVID-19

@@

## Other frequencies

@@

<!-- \appendix -->
<!-- \newpage -->

<!-- # Coefficients, gain and phase shift functions {#sec-an-cof} -->

<!-- ```{r, out.width="100%"} -->
<!-- #| label: fig-graphs-coef-lc -->
<!-- #| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$. -->
<!-- img <- sprintf("img/filters_used/%s.%s", tolower("LC"), "pdf") -->
<!-- knitr::include_graphics(img) -->
<!-- ``` -->

<!-- ```{r, out.width="100%"} -->
<!-- #| label: fig-graphs-coef-ql -->
<!-- #| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$. -->
<!-- img <- sprintf("img/filters_used/%s.%s", tolower("QL"), "pdf") -->
<!-- knitr::include_graphics(img) -->
<!-- ``` -->

<!-- ```{r, out.width="100%"} -->
<!-- #| label: fig-graphs-coef-cq -->
<!-- #| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$. -->
<!-- img <- sprintf("img/filters_used/%s.%s", tolower("CQ"), "pdf") -->
<!-- knitr::include_graphics(img) -->
<!-- ``` -->

<!-- ```{r, out.width="100%"} -->
<!-- #| label: fig-graphs-coef-daf -->
<!-- #| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF). -->
<!-- img <- sprintf("img/filters_used/%s.%s", tolower("DAF"), "pdf") -->
<!-- knitr::include_graphics(img) -->
<!-- ``` -->

>>>>

In @sec-propMM, we describe the general properties of moving averages and the associated quality criteria.
This allows us to understand the foundations behind the construction of moving averages from local polynomial regressions, as well as those behind the two extensions proposed in this article (@sec-nonparamreg).
In @sec-comparison, all these methods are compared empirically on simulated and real series.

@@

The basic assumption  used in  time series decomposition methods is that the input observed time series, $y_t$ can be decomposed into a signal component $\mu_t$ and a an erratic component $\varepsilon_t$ (called irregular):
$$
f(y_t)=\mu_t+\varepsilon_t
$$
where $f$ denotes some appropriate transformation (usually logarithmic or no transformation).
To simplify further notations, $y_t$ will denote the transformed observed series.
The noise component $\varepsilon_t$ is usually assumed to be a white noise.
Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $\mu_t$ represents the trend and cycle components, here estimated jointly and called trend-cycle $TC_t$.
This component can be locally approximated by a polynomial degree (see @sec-nonparamreg):
$$
\mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
$$
with $\xi_t$ a stochastic process uncorrelated with $\varepsilon_t$.
Even if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], an usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
Thus, the trend-cycle $TC_t$ is considered as deterministic and modelled as a polynomial trend of degree $d$.
The coefficients $(\beta_0,\dots,\beta_k)$ can be estimated by weighted least squares.
The estimate $\hat \beta_0$ provides the trend-cycle estimate $\hat TC_t$ and it can be shown that this equivalent to apply of moving average (see @sec-nonparamreg).
The goal of this paper is to present two new methods to separate the trend-cycle component from the noise component in real-time (i.e. when no future point is known), using moving averages build on local polynomial smoothing techniques.

@@

Supplemental materials shows the coefficients, gain and phase shift functions of the four asymmetric filters.

@@

@@

The simulated series are of length 60 years, which is often not realistic for economic time series.
However, since all the methods are local, the results are not affected by the length of the series.
The length of the series would only have an impact for the identification and the estimation of the ARIMA model.
In this case, since the same data generating process is used during the 60 years, it could be relevant to identify the ARIMA model using all the data.
Even if it would improve the results in term of phase shift (see supplemental material) we prefer to only use the last 12 years to identify and estimate the ARIMA model, to be closer to what would have been done with a real time series.

@@

::: {#fig-graphstpsimul}

```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics("img/simulations/phase_shift_simul.pdf")
```
Lecture: this plots represent box plots of the phase shift: the box represents the interquartile range, the central horizontal line the median the dot the outliers and the whisker the range of the data (if no outlier).\
For simulated series with medium variability (when a symmetric filter of length 13 is relevant), with the linear constant (LC) method, 50% of the turning point are detected with a phase shift of at most 5 months whereas when the LC filter is parametrised locally (using), 50% of the turning point are detected with a phase shift of at most 4 months.
For the direct asymmetric filter (DAF), 25% of the turning are detected with a phase shift of at least 7 months (compares to 5 for the LC method).

Distribution of phase shift on simulated series.
:::

@@

## Discussion
### Outliers and COVID-19

@@

### Other frequencies

@@
